---
title: "Coccurrences"
author: "Jarkko Toivonen"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
#library(pROC)
#library(boot)
base_dir <- "~/Amber/SanguinStats/Resultaten/results"
fsummary <- function(df) df %>% mutate(across(where(is.character), as.factor)) %>% summary()
countries <- c("Australia", "Belgium", "Finland", "Netherlands", "South-Africa", "Average") 
countries2 <- sort(c(countries, "Australia2", "South-Africa2"))
result_path <- "~/Amber/SanguinStats/Resultaten/co-occurrence/"
save_figs <- TRUE
```

## TODO

- Find out why Belgian 5don4hb model has different data than in other countries. [DONE] The rows of file
  prediction.csv are permuted.

## Get Finnish data


In the container, just before fitting the (random forest) model the train and validation data sets are saved to files,
e.g. `/tmp/rf-validate-male.rds`. Before the container is killed, these files can be copies out with the `docker cp` command,
e.g. `docker cp kind_brahmagupta:/tmp/rf-validate-male.rds .`, where `kind_brahmagupta` is the name of the container.

```{r Get Finnish datas}
#These dataframes tell in which directories the validation datas are located
# These are with sample size 10,000
validate_subdirectories <- tribble(
  ~Data_country, ~Model, ~Subdirectory,
  "Netherlands", "2don", "~/Amber/Data/SanguinStats/10k/",
  "Netherlands", "5don", "~/Amber/Data/SanguinStats/10k/",
  "Netherlands", "5don4hb", "~/Amber/Data/SanguinStats/10k/",
)
# These are with full data
full_data_validate_subdirectories <- tribble(
  ~Data_country, ~Model, ~Subdirectory,
  "Netherlands", "2don", "~/Amber/Data/SanguinStats/full/",
  "Netherlands", "5don", "~/Amber/Data/SanguinStats/full/",
  "Netherlands", "5don4hb", "~/Amber/Data/SanguinStats/full/",
)
# Read 10k datas
validate_datas <- validate_subdirectories %>%
  mutate(sex=list(c("male", "female"))) %>% 
  unnest_longer(sex) %>%
  mutate(data_file = file.path(Subdirectory, sprintf("rf-validate-%s.rds", sex))) %>%
  mutate(df = map(data_file, function(filename) readRDS(filename)))
# Read full datas
full_data_validate_datas <- full_data_validate_subdirectories %>%
  mutate(sex=list(c("male", "female"))) %>% 
  unnest_longer(sex) %>%
  mutate(data_file = file.path(Subdirectory, sprintf("rf-validate-%s-full.rds", sex))) %>%
  mutate(df = map(data_file, function(filename) readRDS(filename)))
```


```{r Where the fitted models are stored}
folder_trainedmodels <- "~/Amber/SanguinStats/Resultaten/trained_models"
models <- tribble(~Model_country, ~Model, ~dir, ~file,
                  "Finland", "2don", "FI" ,"rf-models-fi-model1-l-2.zip",
                  "Finland", "5don", "FI" ,"rf-models-fi-model2-l-5.zip" ,
                  "Finland", "5don4hb", "FI" ,"rf-models-fi-model3-l-5-hb4.zip",
                  "Netherlands", "2don", "NL", "svm-and-rf-models-NL-2don.zip",
                  "Netherlands", "5don", "NL", "svm-and-rf-models-NL-5don.zip" ,
                  "Netherlands", "5don4hb", "NL", "svm-and-rf-models-NL-5don4hb.zip",
                  "Belgium", "2don", "BE", "svm-and-rf-models_Belgium_2don.zip",
                  "Belgium", "5don", "BE", "svm-and-rf-models_Belgium_5don.zip",
                  "Belgium", "5don4hb", "BE", "svm-and-rf-models_Belgium_5don4hb.zip",
                  # EU thresholds
                  "Australia", "2don", "AU2", "svm-and-rf-models_Australia_2don.zip",
                  "Australia", "5don", "AU2", "svm-and-rf-models_Australia_5don.zip",
                  "Australia", "5don4hb", "AU2", "svm-and-rf-models_Australia_5don4hb.zip",
                  # Own thresholds
                  "Australia2", "2don", "AU", "svm-and-rf-models_Australia_2don.zip",
                  "Australia2", "5don", "AU", "svm-and-rf-models_Australia_5don.zip",
                  "Australia2", "5don4hb", "AU", "svm-and-rf-models_Australia_5don4hb.zip",
                  # EU thresholds
                  "South Africa", "2don", "SA", "svm-and-rf-models_South_Africa_2don.zip",
                  "South Africa", "5don", "SA", "svm-and-rf-models_South_Africa_5don.zip",
                  "South Africa", "5don4hb", "SA", "svm-and-rf-models_South_Africa_5don4hb.zip",
                  # South Africa's own thresholds
                  "South Africa2", "2don", "SA2", "svm-and-rf-models_South_Africa_2don_Own.zip",
                  "South Africa2", "5don", "SA2", "svm-and-rf-models_South_Africa_5don_Own.zip",
                  "South Africa2", "5don4hb", "SA2", "svm-and-rf-models_South_Africa_5don4hb_Own.zip"
)
models <- models %>% 
  mutate(filename=file.path(folder_trainedmodels, dir, file))
```

```{r Function to read the models}
read_model <- function(zipfilename) {
  temp_dir <- "~/Amber/SanguinStats/Resultaten/trained_models"
  cmd <- sprintf("unzip -o %s -d %s", zipfilename, temp_dir)
  #print(cmd)
  cat(sprintf("Reading models from %s\n", zipfilename))
  system(cmd, ignore.stdout = TRUE)
  male_model <- readRDS(sprintf("%s/rf-fit-male.rds", temp_dir))
  female_model <- readRDS(sprintf("%s/rf-fit-female.rds", temp_dir))
  tibble(sex=c("male", "female"), fit=list(male_model, female_model))
}
```

```{r Read the models}
models <- models %>% 
  mutate(fit = map(filename, read_model)) %>%
  unnest(fit)
```

```{r Function that predicts on new data}
mypredict <- function(fit, validate) {
  p <- predict(fit, newdata=validate)
  score <- predict(fit, newdata = validate, type = "prob")[,"Deferred"]
  # sex   original_label predicted_label  score f1_threshold score_predicted_label original_value predicted_value
  result <- validate %>% 
    mutate(original_label = as.integer(Hb_deferral=="Deferred"),
           predicted_label = as.integer(p == "Deferred"),
           score = score,
           f1_threshold = NA,
           score_predicted_label = NA,
           original_value = Hb,
           predicted_value = NA,
           n=1:n()
    ) %>%
    select(sex, original_label, predicted_label, score, f1_threshold, score_predicted_label, 
           original_value, predicted_value, n)
  result
}
```

Predict on 10,000 donors

```{r Compute the predictions}
recomputed_datas <- validate_datas %>% 
  select(Data_country, Model, sex, validate=df) %>%
  full_join(models %>% select(Model_country, Model, sex, fit), by = join_by("Model", "sex"))
recomputed_datas <- recomputed_datas %>%
  mutate(df = map2(fit, validate, mypredict))
```

Predict on full data.

Full data has different data sizes for different model structures (2don, 5don,5don4hb). Sample the datas so that the sample sizes will be equal.

```{r}
tmp <- full_data_validate_datas %>% mutate(n = map_int(df, nrow)) %>% select(-c(Subdirectory, df, data_file))
tmp
min_n <- min(tmp$n)
cat(sprintf("Downsample all data sets to size %i\n", min_n ))
set.seed(123)
filename <- file.path(result_path, "full_data_validate_datas.rds")
if (!file.exists(filename)) {
  full_data_validate_datas <- full_data_validate_datas %>%
    mutate(df = map(df, function(df) slice_sample(df, n=min_n)))
  saveRDS(full_data_validate_datas, filename)
} else {
  full_data_validate_datas <- readRDS(filename)
}
tmp <- full_data_validate_datas %>% 
  mutate(n = map_int(df, nrow),
         deferrals = map_int(df, function(df) sum(df$Hb_deferral=="Deferred")),
         deferral_rate = deferrals / n) %>%
  select(-c(Subdirectory, df, data_file))
tmp
write_tsv(tmp, file.path(result_path, "data_sizes.tsv"))
```

```{r Compute the predictions on full data}
full_data_recomputed_datas <- full_data_validate_datas %>% 
  select(Data_country, Model, sex, validate=df) %>%
  full_join(models %>% select(Model_country, Model, sex, fit), by = join_by(Model, sex))
full_data_recomputed_datas <- full_data_recomputed_datas %>%
  mutate(df = map2(fit, validate, mypredict))
```

## Get the deferral probabilities from the result files

This uses still the old way of using the prediction.csv files outputted by the container.

```{r, eval=FALSE}
folder_results <- "~/Amber/SanguinStats/Resultaten/results"
subdirectories <- tribble(
  ~Model_country, ~Model, ~Subdirectory,
  "Belgium", "2don", "Model1-BE-NL",
  "Belgium", "5don", "Model2-BE-NL",
  "Belgium", "5don4hb", "Model3-BE-NL",
  "South-Africa", "2don", "Model1-SA-NL",
  "South-Africa", "5don", "Model2-SA-NL",
  "South-Africa", "5don4hb", "Model3-SA-NL",
    "South-Africa2", "2don", "Model1-SA2-NL",
  "South-Africa2", "5don", "Model2-SA2-NL",
  "South-Africa2", "5don4hb", "Model3-SA2-NL",
  "Australia", "2don", "Model1-AU-NL",
  "Australia", "5don", "Model2-AU-NL",
  "Australia", "5don4hb", "Model3-AU-NL",
    "Australia2", "2don", "Model1-AU2-NL",
  "Australia2", "5don", "Model2-AU2-NL",
  "Australia2", "5don4hb", "Model3-AU2-NL",
  "Netherlands", "2don", "Model1-NL-NL",
  "Netherlands", "5don", "Model2-NL-NL",
  "Netherlands", "5don4hb", "Model3-NL-NL",
  "Finland", "2don", "Model1-FI-NL",
  "Finland", "5don", "Model2-FI-NL",
  "Finland", "5don4hb", "Model3-FI-NL"
)
```


Get the predictions. Note that it is not obvious that the same observation are sampled and that they are in the same order. For instance, here the results from the Belgian 5don4hb model doesn't seem to match the results from the other models. It would be
better to fix the data set and predict the observations again with different models using the fixed data.
In file docker-apps-plumber.R I permute the rows of the prediction.csv file, hence the mismatch of the rows.

```{r, eval=FALSE}
datas2 <- subdirectories %>%
  mutate(Data_country = "Netherlands",
         prediction_file = file.path(folder_results, Subdirectory, "prediction.csv"),
         df = map(prediction_file, function(filename) read_csv(filename, show_col_types = FALSE)))
```

```{r, eval=FALSE}
datas <- datas2 %>% select(-c(Subdirectory, prediction_file))
```

Check that all result files have same number of rows.

```{r, eval=FALSE}
map_int(datas$df, nrow)
number_of_observations <- datas %>% pluck("df", 1) %>% nrow() / 2
number_of_observations
```

Sample size was 10 000 and the validate set size was 1600. This number of observations are available for both sexes, hence 3200.


## Co-occurrence

Which models to show in plots.

```{r}
models <- c("5don4hb")
#models <- c("5don4hb")
width=70
height=120
```

Combine predictions of each donation attempt by each model.

```{r}
df_with_pred <- recomputed_datas %>% 
  select(-sex) %>% 
  unnest(df) %>% 
  select(-c(validate, fit))
full_data_df_with_pred <- full_data_recomputed_datas %>% 
  select(-sex) %>% 
  unnest(df) %>% 
  select(-c(validate, fit))
#df_with_pred %>% select(Model_country, predicted_label, sex, Model) %>% table()
```


### True predicted deferrals


```{r}
make_cooccurrence_plot <- function(df, title=NULL, original_label=1, predicted_label.x=1, predicted_label.y=1,
                                   color_by_percentage=FALSE) {
  if (color_by_percentage) {
    fill_name <- "percentage" 
    legend.position <- "bottom"
  } else {
    fill_name <- "n"
    legend.position <- "none"
  }
  same_labels_predicted <- is.null(original_label) && is.null(predicted_label.x) && is.null(predicted_label.y)
  symmetric <- same_labels_predicted || predicted_label.x == predicted_label.y  # The count matrices will be symmetric
  tmp <- df %>% 
    select(Model, Model_country, sex, original_label, predicted_label, n) %>%
    filter(!Model_country %in% c("Australia2", "South Africa2"))  # Don't show results for countries' own cutoffs
  countries <- unique(tmp$Model_country)
  models <- unique(tmp$Model)
  number_of_countries <- 5
  number_of_models <- length(models)
  cross_product <- inner_join(tmp, tmp, by=c("Model", "sex", "original_label", "n"), 
                              relationship = "many-to-many") %>%
    relocate(Model, sex, original_label, n)
  if (same_labels_predicted) {
    cooccurrence <- cross_product %>% 
      filter(predicted_label.x == predicted_label.y) %>%
      count(Model, sex, Model_country.x, Model_country.y)
  } else {
    cooccurrence <- cross_product %>% 
      filter(original_label=={{original_label}}, predicted_label.x=={{predicted_label.x}},
             predicted_label.y=={{predicted_label.y}}) %>%
      count(Model, sex, Model_country.x, Model_country.y)
  }
  number_of_observations <- max(df$n)
  number_of_deferrals <- df %>% 
    group_by(sex, Model, Model_country) %>% 
    summarise(deferrals = sum(original_label), .groups="drop") %>%
    select(-Model_country) %>%
    distinct()
  df2 <<- cooccurrence %>%   # Compute success ratio
    left_join(number_of_deferrals, by=c("sex", "Model")) %>%
    mutate(percentage = 100*n/deferrals)
  df2 <<- df2 %>%
    mutate(Model_country.y = fct_rev(Model_country.y), Model_country.x = as.factor(Model_country.x)) %>% 
    mutate(xi = as.numeric(Model_country.x), yi=as.numeric(Model_country.y)) 
  
  if (symmetric) df2 <<- df2 %>%
    filter(xi + yi >= (number_of_countries+1))   # Take only the upper triangle and diagonal
  # There is some aligment/NA problem in geom_raster, hence I explicitly insert one NA cell.
  existing <- df2 %>% select(Model, Model_country.x, Model_country.y, sex) %>% distinct()
  all <- expand_grid(Model=models, sex=c("male", "female"), 
                     Model_country.x=countries, Model_country.y=countries, n=NA, percentage=NA)
  missing <- all %>% anti_join(existing, by=c("Model", "sex", "Model_country.x", "Model_country.y"))
  g <- df2 %>%
    bind_rows(missing) %>%
    ggplot(aes(x=Model_country.x, y=Model_country.y, fill=.data[[fill_name]])) + 
    geom_raster(na.rm=TRUE) +
    geom_text(aes(label=n), color="white", size=3, na.rm=TRUE) +
    scale_y_discrete(limits=rev)+
    scale_fill_continuous(na.value="transparent") +
    labs(x="Model x country", y="Model y country", 
         title=title, 
         #subtitle=sprintf("n=%i", number_of_observations), 
         fill="Percentage of deferrals predicted correctly") +
    (if (number_of_models > 1) facet_grid(sex ~ Model) else facet_grid(sex ~ .)) +
    theme_gray(base_size = 11) +
    theme(legend.position=legend.position,
          axis.text.x = element_text(angle = 45, vjust=1, hjust=1),
          legend.text = element_text(angle = 45, vjust=1, hjust=1))
  g
}


```

### Same label predicted

```{r}
g <- make_cooccurrence_plot(df_with_pred %>% filter(Model %in% models),
                            original_label=NULL, predicted_label.x=NULL, predicted_label.y=NULL, 
                            title="Number of attempts \nlabeled the same way \nin Dutch data")
g
if (save_figs) ggsave(sprintf("%s/same-labels-predicted-10000.png", result_path), g, width=width, height=height, unit="mm")
```

### Agreement

10,000 donors

```{r}
g <- make_cooccurrence_plot(df_with_pred %>% filter(Model %in% models), 
                            title="Number of true deferrals\nthat both models detect\nin Dutch data")
g
if (save_figs) ggsave(sprintf("%s/cooccurrences-of-true-positives-10000.png", result_path), g, width=width, height=height, unit="mm")
```

Full data

```{r}
g <- make_cooccurrence_plot(full_data_df_with_pred %>% filter(Model %in% models), 
                            title="Number of true deferrals\nboth models detect\nin Dutch data")
g
if (save_figs) ggsave(sprintf("%s/cooccurrences-of-true-positives-full-data.png", result_path), g, width=width, height=height, unit="mm")
```


### Disagreement

10,000 donors

```{r}
g <- make_cooccurrence_plot(df_with_pred %>% filter(Model %in% models),
                            title="Number of true deferrals\nmodel x detects in Dutch data \nbut model y does not",
                            original_label=1, predicted_label.x=1, predicted_label.y=0)
g
if (save_figs) ggsave(sprintf("%s/x-better-10000.png", result_path), g, width=width, height=height, unit="mm")
```

Full data

```{r}
g <- make_cooccurrence_plot(full_data_df_with_pred %>% filter(Model %in% models), 
                            title="Number of true deferrals model x detects in\nDutch data but model y does not",
                            original_label=1, predicted_label.x=1, predicted_label.y=0)
g# + coord_fixed()
if (save_figs) ggsave(sprintf("%s/x-better-full-data.pdf", fig_path), g, width=width, height=height, unit="mm")
```

