---
title: "Model averaging"
author: "Jarkko Toivonen"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(pROC)
library(boot)
base_dir <- "~/FRCBS/Hb_predictor_container/results"
fsummary <- function(df) df %>% mutate(across(where(is.character), as.factor)) %>% summary()
countries <- c("Australia", "Belgium", "Finland", "Netherlands", "South-Africa", "Average") 
countries2 <- sort(c(countries, "Australia2", "South-Africa2"))
result_path <- "../results"
fig_path <- file.path(result_path, "figures/pdf")
```

## TODO

- Find out why Belgian 5don4hb model has different data than in other countries. [DONE] The rows of file
  prediction.csv are permuted.

## Get Finnish data


In the container, just before fitting the (random forest) model the train and validation data sets are saved to files,
e.g. `/tmp/rf-validate-male.rds`. Before the container is killed, these files can be copies out with the `docker cp` command,
e.g. `docker cp kind_brahmagupta:/tmp/rf-validate-male.rds .`, where `kind_brahmagupta` is the name of the container.

```{r Get Finnish datas}
# These are with sample size 10,000
validate_subdirectories <- tribble(
  ~Data_country, ~Model, ~Subdirectory,
  "Finland", "2don", "jarkko-2023-03-28-get-data-Australia_2don",
  "Finland", "5don", "jarkko-2023-03-28-get-data-Australia_5don",
  "Finland", "5don4hb", "jarkko-2023-03-28-get-data-Australia_5don4hb",
)
# These are with full data
full_data_validate_subdirectories <- tribble(
  ~Data_country, ~Model, ~Subdirectory,
  "Finland", "2don", "jarkko-2023-08-31-get-data-Australia_full_data_2don",
  "Finland", "5don", "jarkko-2023-08-31-get-data-Australia_full_data_5don",
  "Finland", "5don4hb", "jarkko-2023-08-31-get-data-Australia_full_data_5don4hb",
)
validate_datas <- validate_subdirectories %>%
  mutate(sex=list(c("male", "female"))) %>% 
  unnest_longer(sex) %>%
  mutate(data_file = file.path(base_dir, Subdirectory, sprintf("rf-validate-%s.rds", sex))) %>%
  mutate(df = map(data_file, function(filename) readRDS(filename)))
full_data_validate_datas <- full_data_validate_subdirectories %>%
  mutate(sex=list(c("male", "female"))) %>% 
  unnest_longer(sex) %>%
  mutate(data_file = file.path(base_dir, Subdirectory, sprintf("rf-validate-%s.rds", sex))) %>%
  mutate(df = map(data_file, function(filename) readRDS(filename)))
```


```{r Where the fitted models are stored}
models <- tribble(~Model_country, ~Model, ~dir, ~file,
                  "Finland", "2don", "jarkko-2023-01-17-model1-l-2" ,"svm-and-rf-models.zip",
                  "Finland", "5don", "jarkko-2023-01-17-model2-l-5" ,"svm-and-rf-models.zip",
                  "Finland", "5don4hb", "jarkko-2023-01-17-model3-l-5-hb4" ,"svm-and-rf-models.zip",
"Netherlands", "2don", "amber-2023-01-31", "svm-and-rf-models-2don.zip",
"Netherlands", "5don", "amber-2023-01-31", "svm-and-rf-models-5don.zip",
"Netherlands", "5don4hb", "amber-2023-01-31", "svm-and-rf-models-5don4hb.zip",
"Belgium", "2don", "belgium-2023-03-24", "svm-and-rf-models_Belgium_2don.zip",
"Belgium", "5don", "belgium-2023-03-24", "svm-and-rf-models_Belgium_5don.zip",
"Belgium", "5don4hb", "belgium-2023-03-24", "svm-and-rf-models_Belgium_5don4hb.zip",
# EU thresholds
"Australia", "2don", "australia-2023-03-28-unified_threshold", "svm-and-rf-models_Australia_2don.zip",
"Australia", "5don", "australia-2023-03-28-unified_threshold", "svm-and-rf-models_Australia_5don.zip",
"Australia", "5don4hb", "australia-2023-03-28-unified_threshold", "svm-and-rf-models_Australia_5don4hb.zip",
# Own thresholds
"Australia2", "2don", "australia-2023-02-24", "svm-and-rf-models_Australia_2don.zip",
"Australia2", "5don", "australia-2023-02-24", "svm-and-rf-models_Australia_5don.zip",
"Australia2", "5don4hb", "australia-2023-02-24", "svm-and-rf-models_Australia_5don4hb.zip",
# EU thresholds
"South-Africa", "2don", "south-africa-2023-02-28", "svm-and-rf-models_South_Africa_2don.zip",
"South-Africa", "5don", "south-africa-2023-02-28", "svm-and-rf-models_South_Africa_5don.zip",
"South-Africa", "5don4hb", "south-africa-2023-02-28", "svm-and-rf-models_South_Africa_5don4hb.zip",
# South Africa's own thresholds
"South-Africa2", "2don", "south-africa-2023-04-05-unified_threshold", "svm-and-rf-models_South_Africa_2don_Own.zip",
"South-Africa2", "5don", "south-africa-2023-04-05-unified_threshold", "svm-and-rf-models_South_Africa_5don_Own.zip",
"South-Africa2", "5don4hb", "south-africa-2023-04-05-unified_threshold", "svm-and-rf-models_South_Africa_5don4hb_Own.zip"
)
models <- models %>% 
  mutate(filename=file.path(base_dir, dir, file))
```

```{r Function to read the models}
helper <- function(zipfilename) {
  temp_dir <- tempdir()
  cmd <- sprintf("unzip -o %s -d %s", zipfilename, temp_dir)
  print(cmd)
  system(cmd)
  male_model <- readRDS(sprintf("%s/rf-fit-male.rds", temp_dir))
  female_model <- readRDS(sprintf("%s/rf-fit-female.rds", temp_dir))
  tibble(sex=c("male", "female"), fit=list(male_model, female_model))
}
```

```{r Read the models, echo=FALSE}
models <- models %>% 
  mutate(fit = map(filename, helper)) %>%
  unnest(fit)
```

```{r Function that predicts on new data}
helper2 <- function(fit, validate) {
  p <- predict(fit, newdata=validate)
  score <- predict(fit, newdata = validate, type = "prob")[,"Deferred"]
  # sex   original_label predicted_label  score f1_threshold score_predicted_label original_value predicted_value
  result <- validate %>% 
    mutate(original_label = as.integer(Hb_deferral=="Deferred"),
           predicted_label = as.integer(p == "Deferred"),
           score = score,
           f1_threshold = NA,
           score_predicted_label = NA,
           original_value = Hb,
           predicted_value = NA,
           n=1:n()
    ) %>%
    select(sex, original_label, predicted_label, score, f1_threshold, score_predicted_label, 
           original_value, predicted_value, n)
  result
}
```

```{r Compute the predictions}
# 10,000 donors
recomputed_datas <- validate_datas %>% 
  select(Data_country, Model, sex, validate=df) %>%
  full_join(models %>% select(Model_country, Model, sex, fit))
recomputed_datas <- recomputed_datas %>%
  mutate(df = map2(fit, validate, helper2))
```

Full data has different data sizes. Sample the datas so that the sample sizes will be equal.

```{r}
tmp <- full_data_validate_datas %>% mutate(n = map_int(df, nrow)) %>% select(-c(Subdirectory, df, data_file))
tmp
min_n <- min(tmp$n)
cat(sprintf("Downsample all data sets to size %i\n", min_n ))
full_data_validate_datas <- full_data_validate_datas %>%
  mutate(df = map(df, function(df) slice_sample(df, n=min_n)))
tmp <- full_data_validate_datas %>% mutate(n = map_int(df, nrow)) %>% select(-c(Subdirectory, df, data_file))
tmp
```

```{r Compute the predictions on full data}
full_data_recomputed_datas <- full_data_validate_datas %>% 
  select(Data_country, Model, sex, validate=df) %>%
  full_join(models %>% select(Model_country, Model, sex, fit))
full_data_recomputed_datas <- full_data_recomputed_datas %>%
  mutate(df = map2(fit, validate, helper2))
```

## Get the deferral probabilities from the result files

This uses still the old way of using the prediction.csv files outputted by the container.

```{r, eval=FALSE}
subdirectories <- tribble(
  ~Model_country, ~Model, ~Subdirectory,
  "Belgium", "2don", "jarkko-2023-03-24-Belgium_2don",
  "Belgium", "5don", "jarkko-2023-03-24-Belgium_5don",
  "Belgium", "5don4hb", "jarkko-2023-03-24-Belgium_5don4hb",
  "South-Africa", "2don", "jarkko-2023-02-28-South_Africa_2don",
  "South-Africa", "5don", "jarkko-2023-02-28-South_Africa_5don",
  "South-Africa", "5don4hb", "jarkko-2023-02-28-South_Africa_5don4hb",
  "Australia", "2don", "jarkko-2023-02-27-Australia_2don",
  "Australia", "5don", "jarkko-2023-02-27-Australia_5don",
  "Australia", "5don4hb", "jarkko-2023-02-27-Australia_5don4hb",
  "Netherlands", "2don", "jarkko-2023-01-31-2don",
  "Netherlands", "5don", "jarkko-2023-01-31-5don",
  "Netherlands", "5don4hb", "jarkko-2023-01-31-5don4hb",
  "Finland", "2don", "jarkko-2023-01-17-model1-l-2",
  "Finland", "5don", "jarkko-2023-01-17-model2-l-5",
  "Finland", "5don4hb", "jarkko-2023-01-17-model3-l-5-hb4"
)
```


Get the predictions. Note that it is not obvious that the same observation are sampled and that they are in the same order. For instance, here the results from the Belgian 5don4hb model doesn't seem to match the results from the other models. It would be
better to fix the data set and predict the observations again with different models using the fixed data.
In file docker-apps-plumber.R I permute the rows of the prediction.csv file, hence the mismatch of the rows.

```{r, eval=FALSE}
datas2 <- subdirectories %>%
  mutate(Data_country = "Finland",
         prediction_file = file.path(base_dir, Subdirectory, "prediction.csv"),
         df = map(prediction_file, function(filename) read_csv(filename, show_col_types = FALSE)))
```

```{r, eval=FALSE}
datas <- datas2 %>% select(-c(Subdirectory, prediction_file))
```

Check that all result files have same number of rows.

```{r, eval=FALSE}
map_int(datas$df, nrow)
number_of_observations <- datas %>% pluck("df", 1) %>% nrow() / 2
number_of_observations
```

Sample size was 10 000 and the validate set size was 1600. This number of observations are available for both sexes, hence 3200.

## Average models

### Compute average

```{r}
df <- datas %>% 
  unnest(cols=df) %>% 
  select(-c(id, model, predicted_value, f1_threshold, predicted_label, score_predicted_label)) %>% 
  filter(!(Model_country == "Belgium" & Model == "5don4hb")) %>%  # This is a problematic case
  group_by(Model_country, Model, sex) %>% 
  mutate(n=1:n()) %>%   # Unique identifier for each observation
  ungroup() 
df <- df %>% 
  pivot_wider(names_from=Model_country, values_from=score, 
              id_cols=c(sex, n, Model, original_value, original_label)) #%>%
  #drop_na()
df <- df %>% 
  rowwise() %>% 
  mutate(Average = mean(c_across(any_of(countries)), na.rm=TRUE)) %>% 
  ungroup()
df
summary(df)
```




```{r}
df2 <- df %>% pivot_longer(cols = all_of(countries), 
                           names_to = "Model_country",
                           values_to = "score")
```

### ROC

```{r ROC function}
create_roc_new <- function(labels, score, boot.n=2000) {
  message("Computing the ROC curve")
  tryCatch(error = function(cnd) {
    t <- table(labels, useNA = "always")
    #if (is.null(names(t))) {
      mynames <- c("Accepted (0)", "Deferred (1)")
    #} else mynames <- names(t)
    s <- paste(sprintf("%s: %i", mynames, t), collapse=", ")   # Show the distribution of factor levels in the error message
    cnd$message <- paste("\nThe distribution of response levels is:", s, cnd$message, sep="\n")
    stop(cnd)
  },
           roc <- pROC::roc(response = labels,
                            predictor = score,
                            #smoothed = TRUE,
                            auc = TRUE,
                            legacy.axes = TRUE,   # x-axis is False positive rate instead of specificity
                            xlab = "False Positive Rate", ylab = "True Positive Rate",
                            #percent = TRUE,
                            # arguments for ci
                            ci=TRUE, 
                            conf.level=0.95, 
                            boot.stratified=TRUE,
                            boot.n=boot.n,
                            # arguments for plot
                            plot=FALSE, #plot=TRUE, 
                            main="Receiver operating characteric",
                            #auc.polygon=TRUE, 
                            max.auc.polygon=TRUE, 
                            #grid=TRUE,
                            print.auc=TRUE 
                            #show.thres=FALSE
           )
  )
  AUC <- roc$auc
  #title <- "Receiver operating characteristic"
  title <- "ROC"
  #title <- sprintf("Receiver operating characteristicc (AUC=%.3f)", AUC)
  c <- as.numeric(roc$ci)
  ci <- tibble("AUROC value"=c[2], "AUROC low"=c[1], "AUROC high"=c[3])
  roc_plot <- pROC::ggroc(roc, legacy.axes=TRUE) +
    geom_abline(aes(intercept=0, slope=1), color="lightgray") +
    annotate(geom="text", label=sprintf("AUROC: %.2f (%.2f–%.2f)", c[2], c[1], c[3]), x=0.5, y=0.125) +
    labs(title=title, x = "False positive rate", y = "True positive rate")
  return(list(roc_plot=roc_plot, roc=roc, roc_auc=AUC, roc_ci=ci))
}

```

```{r, results='hide', message=FALSE}
# Compute AUROC values
res_auc <- df2 %>% 
  filter(!(Model == "5don4hb" & Model_country=="Belgium")) %>%
  group_by(sex, Model, Model_country) %>% 
  summarise(obj = list(create_roc_new(original_label, score)), .groups="drop") %>%
  # Make roc_plt, roc object, roc_auc and roc_ci as columns
  hoist(obj, "roc_plot", "roc", "roc_auc", "roc_ci", .simplify=FALSE) %>%
  unnest(roc_ci)
  # unnest_longer(obj) %>%     # For some reason I cannot directly use unnest_wider.
  # pivot_wider(names_from=obj_id, values_from = obj) %>% 
  # mutate(across(c(roc_plot, roc, roc_auc, roc_ci), unname))
```


```{r}
g <- res_auc %>%
  mutate(Model_country = factor(Model_country, levels=countries)) %>%
  mutate(Model_country = fct_rev(Model_country)) %>%
  ggplot(aes(y=Model_country, x=`AUROC value`, xmin=`AUROC low`, xmax=`AUROC high`)) + 
  geom_pointrange(position = position_dodge(width=0.5)) +
  lims(x=0:1) +
  facet_grid(sex~Model)
g
filename <- file.path(fig_path, "model_average_finnish_data_auroc.png")
#ggsave(filename, g)
```

### Precision-recall

```{r Precision-recall function}
# It seems that the number of replicates must be at least as high as the number of rows in the dataframe
# https://stat.ethz.ch/pipermail/r-help/2011-February/269006.html
# boot.n is the number of bootstrap replications, if null use as many replications as there are rows in the dataframe
precision_recall_ci <- function(df, method="norm", boot.n=NULL) {
  
  get_aupr <- function(df, indices) {
    df2 <- df[indices,]
    pb$tick()  # update progress bar
    aupr <- PRROC::pr.curve(scores.class0=df2$score, weights.class0=df2$original_label)$auc.davis.goadrich
    return(aupr)
  }
    
  df <- df %>% select(original_label, score)
  if (is.null(boot.n)) {
    boot.n <- nrow(df)
  }
  
  pb <- progress::progress_bar$new(total = boot.n+1)# init progress bar
  pb$tick(0)
  #p <- progress_estimated(n+1)  # init progress bar
  #b <- boot(df, statistic = get_aupr, R=boot.n, sim="ordinary", stype="i", strata=df$original_label, parallel="multicore")#, ncpus=1)
  b <- boot(df, statistic = get_aupr, R=boot.n, sim="ordinary", stype="i", strata=df$original_label, parallel="no")#, ncpus=1)
  ret <- tryCatch(
    error = function(cnd) return(-1),
    {
      result <- boot.ci(b, conf=0.95, type=method)
      var <- recode(method, "norm"="normal", "perc"="percent", "stud"="student")  # The name of the output field is stupidly sometimes not the same as the parameter name
      ci <- if (method=="norm") result[[var]][2:3] else result[[var]][4:5]
      NULL
    })
  if (!is.null(ret) && ret == -1) {
    ci <- C(NA, NA)
  }
  return(list(ci=ci, result=result))
}
  
create_precision_recall_new <- function(original_label, score, method="norm", boot.n=2000) {
  debug <- TRUE
  message("Computing the precision-recall curve")
  pr_model     <- PRROC::pr.curve(scores.class0=score, weights.class0=original_label, curve=TRUE, rand.compute=TRUE)
  if (debug) message("hep1")
  points <- data.frame(pr_model$curve)
  if (debug) message("hep2")
  AUPR <- pr_model$auc.davis.goadrich
  df <- tibble(original_label=original_label, score=score)
  if (debug) message("hep3")
  prci <- precision_recall_ci(df, method=method, boot.n=boot.n)$ci
  if (debug) message("tassa1")
  title <- "Precision-recall"
  m <- mean(original_label)
  if (debug) message("tassa2")
  pr_plot <- ggplot(points, aes(x=X1,y=X2)) +
    geom_hline(aes(yintercept=m), color="lightgray") +   # theoretical PR curve of random classifier
    annotate(geom="text", label=sprintf("y=%.2f", m), x=0.25, y=m, vjust=-1) +
    geom_line() +
    annotate(geom="text", label=sprintf("AUPR: %.2f (%.2f–%.2f)", AUPR, prci[1], prci[2]), x=0.5, y=0.875) +
    scale_y_continuous(limits=c(0.0, 1.0)) +
    labs(x="Recall",y="Precision", title=title)
  if (debug) message("tassa3")
  ci <- tibble("AUPR value"=AUPR, "AUPR low"=prci[1], "AUPR high"=prci[2])
  if (debug) message("tassa4")
  return(list(pr_plot=pr_plot, pr=points, pr_auc=AUPR, pr_ci=ci))
}

```


```{r, results='hide', message=FALSE}
# Compute AUPR values
res_pr <- df2 %>% 
  filter(!(Model == "5don4hb" & Model_country=="Belgium")) %>%
  group_by(sex, Model, Model_country) %>% 
  summarise(obj = list(create_precision_recall_new(original_label, score)), .groups="drop") %>%
  # Make pr_plt, pr object, pr_auc and pr_ci as columns
  hoist(obj, "pr_plot", "pr", "pr_auc", "pr_ci", .simplify=FALSE) %>%
  unnest(pr_ci)
```


```{r}
g <- res_pr %>%
  mutate(Model_country = factor(Model_country, levels=countries)) %>%
  mutate(Model_country = fct_rev(Model_country)) %>%
  ggplot(aes(y=Model_country, x=`AUPR value`, xmin=`AUPR low`, xmax=`AUPR high`)) + 
  geom_pointrange(position = position_dodge(width=0.5)) + 
  lims(x=0:1) +
  facet_grid(sex~Model)
g
filename <- file.path(fig_path, "model_average_finnish_data_aupr.png")
#ggsave(filename, g)
```

## Co-occurrence

How often models predict each label?


```{r}
# df_with_pred <- df2 %>% 
#   mutate(predicted_label = as.integer(score >= 0.5))
# 10,000 donors
df_with_pred2 <- recomputed_datas %>% 
  select(-sex) %>% 
  unnest(df) %>% 
  select(-c(validate, fit))
full_data_df_with_pred2 <- full_data_recomputed_datas %>% 
  select(-sex) %>% 
  unnest(df) %>% 
  select(-c(validate, fit))
df_with_pred2 %>% select(Model_country, predicted_label, sex, Model) %>% table()
```

### Identically predicted labels

```{r}
cooccurrence <- df_with_pred2 %>% 
  group_by(Model, sex, n, predicted_label) %>%
  summarise(x = list(expand.grid(Model_country, Model_country)), .groups="drop") %>% 
  unnest(x) %>% 
  count(Model, sex, Var1, Var2) %>%
  mutate(across(c("Var1", "Var2"), 
                function(x) fct_relevel(x, setdiff(countries2, "Average")))
  )
```


```{r}
cooccurrence %>%
  filter(Var1 != "Average" & Var2 != "Average") %>%
  mutate(Var2 = fct_rev(Var2)) %>%
  ggplot(aes(x=Var1, y=Var2, fill=n)) + 
  geom_raster() +
  geom_text(aes(label=n), color="white", size=3) +
  labs(x="Model country", y="Model country", title="How often two countries predict the same label", 
       subtitle="Finnish data", fill="Count") +
  facet_grid(sex ~ Model) +
  theme(legend.position="bottom",
        axis.text.x = element_text(angle = 45, vjust=1, hjust=1),
        legend.text = element_text(angle = 45, vjust=1, hjust=1))
```

### True predicted deferrals

```{r, eval=FALSE}
cooccurrence2 <- df_with_pred2 %>% 
  filter(Model_country != "Average",
         original_label == 1,
         predicted_label == 1) %>% 
  group_by(Model, sex, n) %>%
  summarise(x = list(expand.grid(Model_country, Model_country)), .groups="drop") %>% 
  unnest(x) %>% 
  count(Model, sex, Var1, Var2) %>%
  mutate(across(c("Var1", "Var2"), function(x) fct_relevel(x, setdiff(countries2, "Average"))))
```

Another way to do the same in a more general way.

```{r}
make_cooccurrence_plot <- function(df, title=NULL, original_label=1, predicted_label.x=1, predicted_label.y=1,
                                   color_by_percentage=FALSE) {
  if (color_by_percentage) {
    fill_name <- "percentage" 
    legend.position <- "bottom"
  } else {
    fill_name <- "n"
    legend.position <- "nono"
  }
  tmp <- df %>% select(Model, Model_country, sex, original_label, predicted_label, n) %>%
    filter(!Model_country %in% c("Australia2", "South-Africa2"))  # Don't show results for countries' own cutoffs
  cross_product <- inner_join(tmp, tmp, by=c("Model", "sex", "original_label", "n"), 
                              relationship = "many-to-many") %>%
    relocate(Model, sex, original_label, n)
  cooccurrence <- cross_product %>% 
    filter(original_label=={{original_label}}, predicted_label.x=={{predicted_label.x}},
           predicted_label.y=={{predicted_label.y}}) %>%
    count(Model, sex, Model_country.x, Model_country.y)
  number_of_observations <- max(df$n)
  number_of_deferrals <- df %>% 
    group_by(sex, Model, Model_country) %>% summarise(deferrals = sum(original_label), .groups="drop") %>%
    select(-Model_country) %>%
    distinct()
  # There is some aligment/NA problem in geom_raster, hence I explicitly insert one NA cell.
  #tmp <- expand_grid(Model="5don4hb", sex=c("female", "male"), Var1=factor("Belgium"), Var2=factor("Belgium"), n=NA)
  g <- cooccurrence %>%
    #bind_rows(tmp) %>%
    left_join(number_of_deferrals, by=c("sex", "Model")) %>%
    mutate(percentage = 100*n/deferrals) %>%
    #filter(Var1 != "Average" & Var2 != "Average") %>%
    mutate(Model_country.y = fct_rev(Model_country.y)) %>%
    ggplot(aes(x=Model_country.x, y=Model_country.y, fill=.data[[fill_name]])) + 
    geom_raster(na.rm=TRUE) +
    geom_text(aes(label=n), color="white", size=3, na.rm=TRUE) +
    labs(x="Model country", y="Model country", title=title, 
         subtitle=sprintf("Finnish data, n=%i", number_of_observations), 
         fill="Percentage of deferrals predicted correctly") +
    facet_grid(sex ~ Model) +
    theme(legend.position=legend.position,
          axis.text.x = element_text(angle = 45, vjust=1, hjust=1),
          legend.text = element_text(angle = 45, vjust=1, hjust=1))
  g
}
```

10,000 donors
```{r}
g <- make_cooccurrence_plot(df_with_pred2, title="Number of deferrals both models detect")
g
ggsave(sprintf("%s/cooccurrences-of-true-positives-10000.pdf", fig_path), g, width=180, height=120, unit="mm")
```

Full data

```{r}
g <- make_cooccurrence_plot(full_data_df_with_pred2, 
                            title="Number of deferrals both countries detect")
g
ggsave(sprintf("%s/cooccurrences-of-true-positives-full-data.pdf", fig_path), g, width=180, height=120, unit="mm")
```


10,000 donors

```{r}
g <- make_cooccurrence_plot(df_with_pred2, title="Number of deferrals x model detects but y model does not",
                            original_label=1, predicted_label.x=1, predicted_label.y=0)
g
ggsave(sprintf("%s/x-better-10000.pdf", fig_path), g, width=180, height=120, unit="mm")
```

Full data

```{r}
g <- make_cooccurrence_plot(full_data_df_with_pred2, title="Number of deferrals x model detects but y model does not",
                            original_label=1, predicted_label.x=1, predicted_label.y=0)
g
ggsave(sprintf("%s/x-better-full-data.pdf", fig_path), g, width=180, height=120, unit="mm")
```

# old stuff

```{r}
df_with_pred2 %>%
  filter(!(sex == "male" & original_value < 125)) %>%
  filter(Model_country != "Average") %>%
  ggplot(aes(original_value, score, color=Model_country)) +
  geom_line() +
  #geom_point() +
  geom_smooth() +
  labs(x="Hb", y="Probability of deferral", title="Finnish data") +
  facet_wrap(sex ~ Model) +
  theme(legend.position="bottom")
```




