---
title: "Model averaging"
author: "Jarkko Toivonen"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(pROC)
library(boot)
base_dir <- "~/FRCBS/Hb_predictor_container/results"
fsummary <- function(df) df %>% mutate(across(where(is.character), as.factor)) %>% summary()
countries <- c("Australia", "Belgium", "Finland", "Netherlands", "South-Africa", "Average") 
result_path <- "../results"
fig_path <- file.path(result_path, "figures")
```

## TODO

- Find out why Belgian 5don4hb model has different data than in other countries.

## Get Finnish data

These are not used yet.

In the container, just before fitting the (random forest) model the train and validation data sets are saved to files,
e.g. `/tmp/rf-validate-male.rds`. Before the container is killed, these files can be copies out with the `docker cp` command,
e.g. `docker cp kind_brahmagupta:/tmp/rf-validate-male.rds .`, where `kind_brahmagupta` is the name of the container.

```{r}
validate_subdirectories <- tribble(
  ~Data_country, ~Model, ~Subdirectory,
  "Finland", "2don", "jarkko-2023-03-28-get-data-Australia_2don",
  "Finland", "5don", "jarkko-2023-03-28-get-data-Australia_5don",
  "Finland", "5don4hb", "jarkko-2023-03-28-get-data-Australia_5don4hb",
)
validate_datas <- validate_subdirectories %>%
  mutate(sex=list(c("male", "female"))) %>% 
  unnest_longer(sex) %>%
  mutate(data_file = file.path(base_dir, Subdirectory, sprintf("rf-validate-%s.rds", sex))) %>%
  mutate(df = map(data_file, function(filename) readRDS(filename)))
```


## Get the deferral probabilities from the result files

```{r}
subdirectories <- tribble(
  ~Model_country, ~Model, ~Subdirectory,
  "Belgium", "2don", "jarkko-2023-03-24-Belgium_2don",
  "Belgium", "5don", "jarkko-2023-03-24-Belgium_5don",
  "Belgium", "5don4hb", "jarkko-2023-03-24-Belgium_5don4hb",
  "South-Africa", "2don", "jarkko-2023-02-28-South_Africa_2don",
  "South-Africa", "5don", "jarkko-2023-02-28-South_Africa_5don",
  "South-Africa", "5don4hb", "jarkko-2023-02-28-South_Africa_5don4hb",
  "Australia", "2don", "jarkko-2023-02-27-Australia_2don",
  "Australia", "5don", "jarkko-2023-02-27-Australia_5don",
  "Australia", "5don4hb", "jarkko-2023-02-27-Australia_5don4hb",
  "Netherlands", "2don", "jarkko-2023-01-31-2don",
  "Netherlands", "5don", "jarkko-2023-01-31-5don",
  "Netherlands", "5don4hb", "jarkko-2023-01-31-5don4hb",
  "Finland", "2don", "jarkko-2023-01-17-model1-l-2",
  "Finland", "5don", "jarkko-2023-01-17-model2-l-5",
  "Finland", "5don4hb", "jarkko-2023-01-17-model3-l-5-hb4"
)
```


Get the predictions. Note that it is not obvious that the same observation are sampled and that they are in the same order. For instance, here the results from the Belgian 5don4hb model doesn't seem to match the results from the other models. It would be
better to fix the data set and predict the observations again with different models using the fixed data.

```{r}
datas2 <- subdirectories %>%
  mutate(Data_country = "Finland",
         prediction_file = file.path(base_dir, Subdirectory, "prediction.csv"),
         df = map(prediction_file, function(filename) read_csv(filename, show_col_types = FALSE)))
```

```{r}
datas <- datas2 %>% select(-c(Subdirectory, prediction_file))
```

Check that all result files have same number of rows.

```{r}
map_int(datas$df, nrow)
```

Sample size was 10 000 and the validate set size was 1600. This number of observations are available for both sexes, hence 3200.

## Average models

### Compute average

```{r}
df <- datas %>% 
  unnest(cols=df) %>% 
  select(-c(id, model, predicted_value, f1_threshold, predicted_label, score_predicted_label)) %>% 
  filter(!(Model_country == "Belgium" & Model == "5don4hb")) %>%  # This is a problematic case
  group_by(Model_country, Model, sex) %>% 
  mutate(n=1:n()) %>%   # Unique identifier for each observation
  ungroup() 
df <- df %>% 
  pivot_wider(names_from=Model_country, values_from=score, 
              id_cols=c(sex, n, Model, original_value, original_label)) #%>%
  #drop_na()
df <- df %>% 
  rowwise() %>% 
  mutate(Average = mean(c_across(any_of(countries)), na.rm=TRUE)) %>% 
  ungroup()
df
summary(df)
```




```{r}
df2 <- df %>% pivot_longer(cols = all_of(countries), 
                           names_to = "Model_country",
                           values_to = "score")
```

### ROC

```{r ROC function}
create_roc_new <- function(labels, score, boot.n=2000) {
  message("Computing the ROC curve")
  tryCatch(error = function(cnd) {
    t <- table(labels, useNA = "always")
    #if (is.null(names(t))) {
      mynames <- c("Accepted (0)", "Deferred (1)")
    #} else mynames <- names(t)
    s <- paste(sprintf("%s: %i", mynames, t), collapse=", ")   # Show the distribution of factor levels in the error message
    cnd$message <- paste("\nThe distribution of response levels is:", s, cnd$message, sep="\n")
    stop(cnd)
  },
           roc <- pROC::roc(response = labels,
                            predictor = score,
                            #smoothed = TRUE,
                            auc = TRUE,
                            legacy.axes = TRUE,   # x-axis is False positive rate instead of specificity
                            xlab = "False Positive Rate", ylab = "True Positive Rate",
                            #percent = TRUE,
                            # arguments for ci
                            ci=TRUE, 
                            conf.level=0.95, 
                            boot.stratified=TRUE,
                            boot.n=boot.n,
                            # arguments for plot
                            plot=FALSE, #plot=TRUE, 
                            main="Receiver operating characteric",
                            #auc.polygon=TRUE, 
                            max.auc.polygon=TRUE, 
                            #grid=TRUE,
                            print.auc=TRUE 
                            #show.thres=FALSE
           )
  )
  AUC <- roc$auc
  #title <- "Receiver operating characteristic"
  title <- "ROC"
  #title <- sprintf("Receiver operating characteristicc (AUC=%.3f)", AUC)
  c <- as.numeric(roc$ci)
  ci <- tibble("AUROC value"=c[2], "AUROC low"=c[1], "AUROC high"=c[3])
  roc_plot <- pROC::ggroc(roc, legacy.axes=TRUE) +
    geom_abline(aes(intercept=0, slope=1), color="lightgray") +
    annotate(geom="text", label=sprintf("AUROC: %.2f (%.2f–%.2f)", c[2], c[1], c[3]), x=0.5, y=0.125) +
    labs(title=title, x = "False positive rate", y = "True positive rate")
  return(list(roc_plot=roc_plot, roc=roc, roc_auc=AUC, roc_ci=ci))
}

```

```{r, results='hide', message=FALSE}
# Compute AUROC values
res_auc <- df2 %>% 
  filter(!(Model == "5don4hb" & Model_country=="Belgium")) %>%
  group_by(sex, Model, Model_country) %>% 
  summarise(obj = list(create_roc_new(original_label, score)), .groups="drop") %>%
  # Make roc_plt, roc object, roc_auc and roc_ci as columns
  hoist(obj, "roc_plot", "roc", "roc_auc", "roc_ci", .simplify=FALSE) %>%
  unnest(roc_ci)
  # unnest_longer(obj) %>%     # For some reason I cannot directly use unnest_wider.
  # pivot_wider(names_from=obj_id, values_from = obj) %>% 
  # mutate(across(c(roc_plot, roc, roc_auc, roc_ci), unname))
```


```{r}
g <- res_auc %>%
  mutate(Model_country = factor(Model_country, levels=countries)) %>%
  mutate(Model_country = fct_rev(Model_country)) %>%
  ggplot(aes(y=Model_country, x=`AUROC value`, xmin=`AUROC low`, xmax=`AUROC high`)) + 
  geom_pointrange(position = position_dodge(width=0.5)) +
  lims(x=0:1) +
  facet_grid(sex~Model)
g
filename <- file.path(fig_path, "model_average_finnish_data_auroc.png")
#ggsave(filename, g)
```

### Precision-recall

```{r Precision-recall function}
# It seems that the number of replicates must be at least as high as the number of rows in the dataframe
# https://stat.ethz.ch/pipermail/r-help/2011-February/269006.html
# boot.n is the number of bootstrap replications, if null use as many replications as there are rows in the dataframe
precision_recall_ci <- function(df, method="norm", boot.n=NULL) {
  
  get_aupr <- function(df, indices) {
    df2 <- df[indices,]
    pb$tick()  # update progress bar
    aupr <- PRROC::pr.curve(scores.class0=df2$score, weights.class0=df2$original_label)$auc.davis.goadrich
    return(aupr)
  }
    
  df <- df %>% select(original_label, score)
  if (is.null(boot.n)) {
    boot.n <- nrow(df)
  }
  
  pb <- progress::progress_bar$new(total = boot.n+1)# init progress bar
  pb$tick(0)
  #p <- progress_estimated(n+1)  # init progress bar
  #b <- boot(df, statistic = get_aupr, R=boot.n, sim="ordinary", stype="i", strata=df$original_label, parallel="multicore")#, ncpus=1)
  b <- boot(df, statistic = get_aupr, R=boot.n, sim="ordinary", stype="i", strata=df$original_label, parallel="no")#, ncpus=1)
  ret <- tryCatch(
    error = function(cnd) return(-1),
    {
      result <- boot.ci(b, conf=0.95, type=method)
      var <- recode(method, "norm"="normal", "perc"="percent", "stud"="student")  # The name of the output field is stupidly sometimes not the same as the parameter name
      ci <- if (method=="norm") result[[var]][2:3] else result[[var]][4:5]
      NULL
    })
  if (!is.null(ret) && ret == -1) {
    ci <- C(NA, NA)
  }
  return(list(ci=ci, result=result))
}
  
create_precision_recall_new <- function(original_label, score, method="norm", boot.n=2000) {
  debug <- TRUE
  message("Computing the precision-recall curve")
  pr_model     <- PRROC::pr.curve(scores.class0=score, weights.class0=original_label, curve=TRUE, rand.compute=TRUE)
  if (debug) message("hep1")
  points <- data.frame(pr_model$curve)
  if (debug) message("hep2")
  AUPR <- pr_model$auc.davis.goadrich
  df <- tibble(original_label=original_label, score=score)
  if (debug) message("hep3")
  prci <- precision_recall_ci(df, method=method, boot.n=boot.n)$ci
  if (debug) message("tassa1")
  title <- "Precision-recall"
  m <- mean(original_label)
  if (debug) message("tassa2")
  pr_plot <- ggplot(points, aes(x=X1,y=X2)) +
    geom_hline(aes(yintercept=m), color="lightgray") +   # theoretical PR curve of random classifier
    annotate(geom="text", label=sprintf("y=%.2f", m), x=0.25, y=m, vjust=-1) +
    geom_line() +
    annotate(geom="text", label=sprintf("AUPR: %.2f (%.2f–%.2f)", AUPR, prci[1], prci[2]), x=0.5, y=0.875) +
    scale_y_continuous(limits=c(0.0, 1.0)) +
    labs(x="Recall",y="Precision", title=title)
  if (debug) message("tassa3")
  ci <- tibble("AUPR value"=AUPR, "AUPR low"=prci[1], "AUPR high"=prci[2])
  if (debug) message("tassa4")
  return(list(pr_plot=pr_plot, pr=points, pr_auc=AUPR, pr_ci=ci))
}

```


```{r, results='hide', message=FALSE}
# Compute AUPR values
res_pr <- df2 %>% 
  filter(!(Model == "5don4hb" & Model_country=="Belgium")) %>%
  group_by(sex, Model, Model_country) %>% 
  summarise(obj = list(create_precision_recall_new(original_label, score)), .groups="drop") %>%
  # Make pr_plt, pr object, pr_auc and pr_ci as columns
  hoist(obj, "pr_plot", "pr", "pr_auc", "pr_ci", .simplify=FALSE) %>%
  unnest(pr_ci)
```


```{r}
g <- res_pr %>%
  mutate(Model_country = factor(Model_country, levels=countries)) %>%
  mutate(Model_country = fct_rev(Model_country)) %>%
  ggplot(aes(y=Model_country, x=`AUPR value`, xmin=`AUPR low`, xmax=`AUPR high`)) + 
  geom_pointrange(position = position_dodge(width=0.5)) + 
  lims(x=0:1) +
  facet_grid(sex~Model)
g
filename <- file.path(fig_path, "model_average_finnish_data_aupr.png")
#ggsave(filename, g)
```

## Co-occurrence

How often models predict each label?

```{r}
df_with_pred <- df2 %>% 
  mutate(predicted_label = as.integer(score >= 0.5))
df_with_pred %>% select(Model_country, predicted_label, sex, Model) %>% table()
```

```{r}
cooccurrence <- df_with_pred %>% 
  #filter(Model=="2don", sex=="female") %>% 
  group_by(Model, sex, n, predicted_label) %>%
  summarise(x = list(expand.grid(Model_country, Model_country)), .groups="drop") %>% 
  unnest(x) %>% 
  count(Model, sex, Var1, Var2) %>%
  mutate(across(c("Var1", "Var2"), function(x) fct_relevel(x, countries)))
```


```{r}
cooccurrence %>%
  filter(Var1 != "Average" & Var2 != "Average") %>%
  mutate(Var2 = fct_rev(Var2)) %>%
  ggplot(aes(x=Var1, y=Var2, fill=n)) + 
  geom_raster() +
  geom_text(aes(label=n), color="white", size=3) +
  labs(x="Model country", y="Model country", title="How often two countries predict the same label", 
       subtitle="Finnish data", fill="Count") +
  facet_grid(sex ~ Model) +
  theme(legend.position="bottom",
        axis.text.x = element_text(angle = 45, vjust=1, hjust=1),
        legend.text = element_text(angle = 45, vjust=1, hjust=1))
```

```{r}
df_with_pred %>%
  filter(!(sex == "male" & original_value < 125)) %>%
  filter(Model_country != "Average") %>%
  ggplot(aes(original_value, score, color=Model_country)) +
  geom_line() +
  #geom_point() +
  geom_smooth() +
  labs(x="Hb", y="Probability of deferral", title="Finnish data") +
  facet_wrap(sex ~ Model) +
  theme(legend.position="bottom")
```




